skeleton

### 2023-10-30 inst/T20.Rmd
add hex sticker to more Rmd files
needs to be a single line for knitr
name all chunks and cache the slow ones

```{r}
library("hyper2",quietly=TRUE)
T20_table <- read.table("T20.txt",header=TRUE)
head(T20_table)
nrow(T20_table)
```

@@

```{r}
x <- table(T20_table$toss_winner==T20_table$match_winner)
x
binom.test(x,alternative="less")$p.value
```

@@

```{r}
table(T20_table$toss_decision)
```

@@

```{r}
attach(T20_table)
bat_first_wins <- 
((toss_winner==match_winner)&(toss_decision=='bat'  )) | 
((toss_winner!=match_winner)&(toss_decision=='field'))
sum(bat_first_wins)
field_first_wins <- 
((toss_winner==match_winner)&(toss_decision=='field'  )) | 
((toss_winner!=match_winner)&(toss_decision=='bat'))
sum(field_first_wins)
detach(T20_table)
```

@@

```{r}
sum(field_first_wins) + sum(bat_first_wins)
nrow(T20_table)
```
agreeing as expected.

@@

```{r}
M <- table(
decision           = T20_table$toss_decision,
won_toss_won_match = T20_table$toss_winner==T20_table$match_winner)
dimnames(M) <- list(
decision=c('bat first','field first'),
won_match =c(FALSE,TRUE))
M
```

We may reject homogoneity of proportion (Fisher's exact test, $p= `r
round(fisher.test(M)$p.value,3)`$).  Recalling that a slight majority
of toss winners elect to field first (361 out of 633), we see that
this is a consistent result.

@@

```{r}
attach(lapply(T20_table[, ], as.vector))  # vector access to team1, etc
T20 <- hyper2()
for(i in seq_along(team1)){
   T20[match_winner[i]] %<>% inc
   T20[c(team1[i],team2[i])] %<>% dec

@@

```{r}
T20_maxp <- maxp(T20)
dotchart(T20_maxp)
```

@@

```{r}
a1 <- maxp(T20,give=TRUE)$value        # likelihood at the evaluate
a2 <- loglik(indep(equalp(T20)),T20)  # likelihood at p1=p2=...=p14
a1-a2
```

@@

```{r}
pchisq(2*14.5,df=13,lower.tail=FALSE)
```

@@

```{r}
T20_toss  <- hyper2(pnames=c('toss',levels(T20_table$team1)))
for(i in seq_along(team1)){
      players <- c(team1[i],team2[i])
      if(toss_winner[i] == match_winner[i]){ # win toss, win match
         T20_toss[c('toss',match_winner[i])] %<>% inc
      } else { # win toss, lose match
         T20_toss[match_winner[i]] %<>% inc
      }			 
   T20_toss[c('toss',players)] %<>% dec
   T20_toss
}			  
```
and again we can maximize the likelihood:

```{r}
T20_maxp_toss <- maxp(T20_toss)
dotchart(T20_maxp_toss)
```

>>>>

```{r kickoff}
library("hyper2",quietly=TRUE)
T20_table <- read.table("T20.txt",header=TRUE)
head(T20_table)
nrow(T20_table)
```

@@

```{r onewayanal}
x <- table(T20_table$toss_winner==T20_table$match_winner)
x
binom.test(x,alternative="less")$p.value
```

@@

```{r showtable}
table(T20_table$toss_decision)
```

@@

```{r pbtw}
attach(T20_table)
bat_first_wins <- 
((toss_winner==match_winner)&(toss_decision=='bat'  )) | 
((toss_winner!=match_winner)&(toss_decision=='field'))
sum(bat_first_wins)
field_first_wins <- 
((toss_winner==match_winner)&(toss_decision=='field'  )) | 
((toss_winner!=match_winner)&(toss_decision=='bat'))
sum(field_first_wins)
detach(T20_table)
```

@@

```{r conscheck}
sum(field_first_wins) + sum(bat_first_wins)
nrow(T20_table)
```
agreeing as expected.

@@

```{r twowaycontables}
M <- table(
decision           = T20_table$toss_decision,
won_toss_won_match = T20_table$toss_winner==T20_table$match_winner)
dimnames(M) <- list(
decision=c('bat first','field first'),
won_match =c(FALSE,TRUE))
M
```

We may reject homogeneity of proportion (Fisher's exact test, $p= `r round(fisher.test(M)$p.value,3)`$).  Recalling that a slight majority
of toss winners elect to field first (361 out of 633), we see that
this is a consistent result.

@@

```{r indstrength}
attach(lapply(T20_table[, ], as.vector))  # vector access to team1, etc
T20 <- hyper2()
for(i in seq_along(team1)){
   T20[match_winner[i]] %<>% inc
   T20[c(team1[i],team2[i])] %<>% dec

@@

```{r maxisstraight,cache=TRUE}
T20_maxp <- maxp(T20)
dotchart(T20_maxp)
```

@@

```{r nullallteams}
a1 <- maxp(T20,give=TRUE)$value        # likelihood at the evaluate
a2 <- loglik(indep(equalp(T20)),T20)  # likelihood at p1=p2=...=p14
a1-a2
```

@@

```{r wilkstest}
pchisq(2*14.5,df=13,lower.tail=FALSE)
```

@@

```{r tossdiff}
T20_toss  <- hyper2(pnames=c('toss',levels(T20_table$team1)))
for(i in seq_along(team1)){
      players <- c(team1[i],team2[i])
      if(toss_winner[i] == match_winner[i]){ # win toss, win match
         T20_toss[c('toss',match_winner[i])] %<>% inc
      } else { # win toss, lose match
         T20_toss[match_winner[i]] %<>% inc
      }			 
   T20_toss[c('toss',players)] %<>% dec
   T20_toss
}			  
```
and again we can maximize the likelihood:

```{r maxlike,cache=TRUE}
T20_maxp_toss <- maxp(T20_toss)
dotchart(T20_maxp_toss)
```

### 2023-10-30 inst/formula1.Rmd
name all chunks and cache the slow ones

```{r,cache=TRUE}
mf <- maxp(formula1)
l0 <- loglik(indep(mf),formula1)
l1 <- loglik(indep(zipf(size(formula1))),formula1)
c(l0,l1)
pchisq(2*(l0-l1),df=size(formula1)-1,lower.tail=FALSE)
```

@@

```{r}
mf
mf[rank(-mf)] <- zipf(size(formula1))
mf
l2 <- loglik(indep(mf),formula1)
l2
pchisq(2*(l0-l1),df=size(formula1)-1,lower.tail=FALSE)
```

@@

```{r,testwetstrengthzero}
pnames(H)
H9 <- keep_flawed(H,c("S",pnames(H)[1:8]))
specificp.gt.test(H9,"S",0)
```

>>>>

```{r twonatural,cache=TRUE}
mf <- maxp(formula1)
l0 <- loglik(indep(mf),formula1)
l1 <- loglik(indep(zipf(size(formula1))),formula1)
c(l0,l1)
pchisq(2*(l0-l1),df=size(formula1)-1,lower.tail=FALSE)
```

@@

```{r ranklike}
mf
mf[rank(-mf)] <- zipf(size(formula1))
mf
l2 <- loglik(indep(mf),formula1)
l2
pchisq(2*(l0-l1),df=size(formula1)-1,lower.tail=FALSE)
```

@@

```{r,testwetstrengthzero,cache=TRUE}
pnames(H)
H9 <- keep_flawed(H,c("S",pnames(H)[1:8]))
specificp.gt.test(H9,"S",0)
```

### 2023-10-30 inst/grad_weight.Rmd
name all chunks and cache the slow ones

```{r}
H <- hyper3()
f <- function(w){
  H <- hyper3()
  H[c(p1=1 , p2=w , p3=w , p4=1)] <- 5
  H
}
w <- 1.4
dw <- 1e-6
HH <- f(w)
HH
probs <- c(p1=0.1,p2=0.2,p3=0.3,p4=0.4)
wpv <- c(p1=1 , p2=w , p3=w , p4=1)
LHS <- grad_weight_single(HH, probs, wpv, c("p2","p3"))
RHS <-  (loglik(probs,f(w=w+dw)) - loglik(probs,f(w=w)))/dw
c(RHS,LHS,LHS-RHS)
```

@@

```{r}

@@

```{r}
diff <- 0
dw <- 1e-4
probs <- c(p1=0.15 , p2=0.45, p3=0.4)
B <- brackfunc(w,D)
for(i in seq_along(wanted)){
   diff <- diff +  grad_weight_single(H, probs, wpv=B[[i]],wanted=wanted[[i]])
}

@@

```{r}
RHS2 <- (loglik(probs,f2(w=w+dw/2,D=D)) - loglik(probs,f2(w=w-dw/2,D=D)))/dw  # differential via numerics
c(LHS,RHS2,LHS-RHS2)
```

@@

```{r}
o <- c("p1","p2")
n <- NULL
wantedD <- list(n,o,n,o,n,n,o,n,n)

>>>>

```{r slightlymorecomplicated}
H <- hyper3()
f <- function(w){
  H <- hyper3()
  H[c(p1=1 , p2=w , p3=w , p4=1)] <- 5
  H
}
w <- 1.4
dw <- 1e-6
HH <- f(w)
HH
probs <- c(p1=0.1,p2=0.2,p3=0.3,p4=0.4)
wpv <- c(p1=1 , p2=w , p3=w , p4=1)
LHS <- grad_weight_single(HH, probs, wpv, c("p2","p3"))
RHS <-  (loglik(probs,f(w=w+dw)) - loglik(probs,f(w=w)))/dw
c(RHS,LHS,LHS-RHS)
```

@@

```{r chessexample}

@@

```{r diffwrtw}
diff <- 0
dw <- 1e-4
probs <- c(p1=0.15 , p2=0.45, p3=0.4)
B <- brackfunc(w,D)
for(i in seq_along(wanted)){
   diff <- diff +  grad_weight_single(H, probs, wpv=B[[i]],wanted=wanted[[i]])
}

@@

```{r cleverusecentral}
RHS2 <- (loglik(probs,f2(w=w+dw/2,D=D)) - loglik(probs,f2(w=w-dw/2,D=D)))/dw  # differential via numerics
c(LHS,RHS2,LHS-RHS2)
```

@@

```{r diffwrtD}
o <- c("p1","p2")
n <- NULL
wantedD <- list(n,o,n,o,n,n,o,n,n)

### 2023-10-30 inst/gradient.Rmd
name all chunks and cache the slow ones

```{r}
chess
```

@@

```{r}
chess_gradient_theoretical <- function(x){
    Topalov <- x[1]
    Anand   <- x[2]
    Karpov  <- x[3]

@@

```{r}
p_chess <- 
	c(
		Topalov = 0.5,
		Anand   = 0.3,
		Karpov  = 0.2
	)
chess_gradient_theoretical(p_chess) - chess_gradient_numerical(p_chess)
```

@@

```{r}
icons
```

@@

```{r}
icons_gradient_theoretical <- function(x){
    NB <- x[1]
    L <- x[2]
    PB <- x[3]
    THC <- x[4]
    OA <- x[5]
    WAIS <- x[6]

@@

```{r}
p_icons <- c(NB=0.3, L=0.1, PB=0.2, THC=0.15, OA=0.15, WAIS=0.1)
```

```{r}
icons_gradient_theoretical(p_icons)
icons_gradient_theoretical(p_icons) - icons_gradient_numerical(p_icons)
```

@@

```{r}
rbind(indep(p_chess), gradient(chess,indep(p_chess)))
rbind(p_chess,gradientn(chess,p_chess))
```

@@

```{r}
sum(powers(chess))
sum(p_chess*gradientn(chess,p_chess))
```

@@

```{r}
sum(powers(icons))
sum(p_icons*gradientn(icons,p_icons))
```

@@

```{r}
(M <- hessian(icons))
is_ok_hessian(M)
```

@@

```{r}
is_ok_hessian(M,give=TRUE)
```

@@

```{r}
(p <- indep(equalp(icons))) # NB not the evaluate
(G <- gradient(icons,p))
(M <- hessian(icons,p,border=FALSE))
eigen(M,TRUE,TRUE)$values
```

```{r}
dp <- rnorm(5)*1e-4        # small perturbation
dL <- loglik(p+dp,icons) - loglik(p,icons)  # delta loglikelihood (exact)

@@

```{r}

>>>>

```{r showchess}
chess
```

@@

```{r chessgrad}
chess_gradient_theoretical <- function(x){
    Topalov <- x[1]
    Anand   <- x[2]
    Karpov  <- x[3]

@@

```{r pchessagree}
p_chess <- 
	c(
		Topalov = 0.5,
		Anand   = 0.3,
		Karpov  = 0.2
	)
chess_gradient_theoretical(p_chess) - chess_gradient_numerical(p_chess)
```

@@

```{r showicons}
icons
```

@@

```{r samethingalgebra}
icons_gradient_theoretical <- function(x){
    NB <- x[1]
    L <- x[2]
    PB <- x[3]
    THC <- x[4]
    OA <- x[5]
    WAIS <- x[6]

@@

```{r isicons}
p_icons <- c(NB=0.3, L=0.1, PB=0.2, THC=0.15, OA=0.15, WAIS=0.1)
```

```{r comparegradientwithwithout}
icons_gradient_theoretical(p_icons)
icons_gradient_theoretical(p_icons) - icons_gradient_numerical(p_icons)
```

@@

```{r lookatchess}
rbind(indep(p_chess), gradient(chess,indep(p_chess)))
rbind(p_chess,gradientn(chess,p_chess))
```

@@

```{r isderivzero}
sum(powers(chess))
sum(p_chess*gradientn(chess,p_chess))
```

@@

```{r seesumpower}
sum(powers(icons))
sum(p_icons*gradientn(icons,p_icons))
```

@@

```{r ishessian}
(M <- hessian(icons))
is_ok_hessian(M)
```

@@

```{r isnothessian}
is_ok_hessian(M,give=TRUE)
```

@@

```{r evaluatehessian}
(p <- indep(equalp(icons))) # NB not the evaluate
(G <- gradient(icons,p))
(M <- hessian(icons,p,border=FALSE))
eigen(M,TRUE,TRUE)$values
```

```{r smallpert}
dp <- rnorm(5)*1e-4        # small perturbation
dL <- loglik(p+dp,icons) - loglik(p,icons)  # delta loglikelihood (exact)

@@

```{r useborderedhess}

### 2023-10-30 inst/karate.Rmd
name all chunks and cache the slow ones

```{r}
karate_maxp <- maxp(karate)
```

```{r}
dotchart(log10(karate_maxp))
```

@@

```{r}
pk <- pnames(karate)
M <- matrix(0,length(pk),length(pk))
rownames(M) <- pk
colnames(M) <- pk
o <- karate_table # saves typing
for(i in seq_len(nrow(o))){
  M[o[i,1],o[i,3]] %<>% `+`(o[i,2])
  M[o[i,3],o[i,1]] %<>% `+`(o[i,4])
}
```

```{r,cache=TRUE}
pM <- pairwise(M)
pM == karate  # consistency check; should be TRUE
```

@@

```{r}
karate_zermelo <- zermelo(M)
loglik(karate_zermelo,karate)
loglik(karate_maxp,karate)
```

@@

```{r}
par(pty='s')
plot(log10(karate_zermelo),log10(karate_maxp),asp=1,xlim=c(-4,0),ylim=c(-4,0))
abline(0,1)
grid()
```

>>>>

```{r  calcmaxp,cache=TRUE}
karate_maxp <- maxp(karate)
```

```{r showdot}
dotchart(log10(karate_maxp))
```

@@

```{r defmatwinloss}
pk <- pnames(karate)
M <- matrix(0,length(pk),length(pk))
rownames(M) <- pk
colnames(M) <- pk
o <- karate_table # saves typing
for(i in seq_len(nrow(o))){
  M[o[i,1],o[i,3]] %<>% `+`(o[i,2])
  M[o[i,3],o[i,1]] %<>% `+`(o[i,4])
}
```

```{r,label=pairwisem,cache=TRUE}
pM <- pairwise(M)
pM == karate  # consistency check; should be TRUE
```

@@

```{r usezerm}
karate_zermelo <- zermelo(M)
loglik(karate_zermelo,karate)
loglik(karate_maxp,karate)
```

@@

```{r plotzerm}
par(pty='s')
plot(log10(karate_zermelo),log10(karate_maxp),asp=1,xlim=c(-4,0),ylim=c(-4,0))
abline(0,1)
grid()
```

### 2023-10-30 inst/kka.Rmd
name all chunks and cache the slow ones

```{r}
l <- l-max(l)
contour(lambda,D,l,nlevels=30,xlab="lambda",ylab="D")
abline(v=lambda,col='gray',lwd=0.2)
abline(h=D,col='gray',lwd=0.2)
lambda
D
```

>>>>

```{r contourplotter}
l <- l-max(l)
contour(lambda,D,l,nlevels=30,xlab="lambda",ylab="D")
abline(v=lambda,col='gray',lwd=0.2)
abline(h=D,col='gray',lwd=0.2)
lambda
D
```

### 2023-10-30 inst/skeleton.Rmd
name all chunks and cache the slow ones

```{r}
bm1 <- st[,1] > st["Maier",1]  # lost to Maier in run 1
bm2 <- st[,2] > st["Maier",2]  # lost to Maier in run 2
bm3 <- st[,3] > st["Maier",3]  # lost to Maier in run 3
bm4 <- st[,4] > st["Maier",4]  # lsot to Maier in run 4
st[bm1 | bm2 | bm3 | bm4,]
```

@@

```{r}
rownames(st[bm1 | bm2 | bm3 | bm4,])
jj <- cbind(won="Maier",lost=rownames(st[bm1 | bm2 | bm3 | bm4,]))
noquote(jj)
plot(graph(c(t(jj))))
```

@@

```{r}
beats <- function(p){   # returns everyone whom competitor 'p' beats
  bm1 <- st[,1] > st[p,1]  
  bm2 <- st[,2] > st[p,2]  
  bm3 <- st[,3] > st[p,3] 
  bm4 <- st[,4] > st[p,4] 
  jj <- cbind(p,rownames(st[bm1 | bm2 | bm3 | bm4,]))
  return(c(t(jj)))
}
bb <- c()
for(i in rownames(st)){bb <- c(bb,beats(i))}
graph(bb)
plot(graph(bb))
```

@@

```{r}
(st2 <- st[-(4:19),])  # exclude a bunch of mid-range competitors
```

@@

```{r}
bb <- c()
for(i in rownames(st2)){bb <- c(bb,beats(i))}
graph(bb)
plot(graph(bb),layout = layout.fruchterman.reingold)
```

@@

```{r}
plot(pvals,log='y',pch=16)
for(i in seq_along(pvals)){text(i,pvals[i],names(pvals)[i],pos=4,col='gray')}
```

@@

```{r,label=savecurlingdataset}
save(skeleton_table,skeleton,skeleton_maxp,file="skeleton.rda")
```

>>>>

```{r identifycompetitors}
bm1 <- st[,1] > st["Maier",1]  # lost to Maier in run 1
bm2 <- st[,2] > st["Maier",2]  # lost to Maier in run 2
bm3 <- st[,3] > st["Maier",3]  # lost to Maier in run 3
bm4 <- st[,4] > st["Maier",4]  # lsot to Maier in run 4
st[bm1 | bm2 | bm3 | bm4,]
```

@@

```{r losttomaier}
rownames(st[bm1 | bm2 | bm3 | bm4,])
jj <- cbind(won="Maier",lost=rownames(st[bm1 | bm2 | bm3 | bm4,]))
noquote(jj)
plot(graph(c(t(jj))))
```

@@

```{r definebeats}
beats <- function(p){   # returns everyone whom competitor 'p' beats
  bm1 <- st[,1] > st[p,1]  
  bm2 <- st[,2] > st[p,2]  
  bm3 <- st[,3] > st[p,3] 
  bm4 <- st[,4] > st[p,4] 
  jj <- cbind(p,rownames(st[bm1 | bm2 | bm3 | bm4,]))
  return(c(t(jj)))
}
bb <- c()
for(i in rownames(st)){bb <- c(bb,beats(i))}
graph(bb)
plot(graph(bb))
```

@@

```{r excludemidrange}
(st2 <- st[-(4:19),])  # exclude a bunch of mid-range competitors
```

@@

```{r justtop3}
bb <- c()
for(i in rownames(st2)){bb <- c(bb,beats(i))}
graph(bb)
plot(graph(bb),layout = layout.fruchterman.reingold)
```

@@

```{r showpvals}
plot(pvals,log='y',pch=16)
for(i in seq_along(pvals)){text(i,pvals[i],names(pvals)[i],pos=4,col='gray')}
```

@@

```{r,label=saveskeletondataset}
save(skeleton_table,skeleton,skeleton_maxp,file="skeleton.rda")
```
